### 2017-12-23 cuthesis_phd.tex
using thesis template for now

\documentstyle[cuthesis]{report}

\author{Valerie Hayot-Sasson}
\title{Thesis Title}
\titleOfPhDAuthor{Ms.}          % or Ms., Mrs., Miss, etc. (only for PhD's)
\PhD                            % Masters by default
\degree{Computer Science}
% \dept{Computer Science}       % default is Comp.Sci.
\cosupervisor                   % if you also have a co-supervisor

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{abstract}
  Text of abstract.  
\end{abstract}

\begin{acknowledgments}
  Text of acknowledgments.
\end{acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Body of Thesis goes here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addcontentsline{toc}{chapter}{Bibliography}
%\bibliography{abbr,chalin,common,larch,tn}  %place your .bib files here
%\bibliographystyle{alpha}                   %the bibliography style to use

\end{document}